The ChEMBL database as Linked Open Data

### 2013-03-06 article.tex
Spell checked

The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
The SQL data dump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
were then executed to generate RDF triples using a set of PHP scripts. The scripts use SQL to
query the database and output triples using print statements. Custom PHP scripts have been used
as a light-weight soluton to provide a Linked Data API via a Apache webserver, as well as to
create data files with triples. These scripts are Open Source 
and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest
Group~\cite{Marshall2012}. This process has not changed much since its original description.

@@

However, where the database content is well-defined, specific ontologies have been used.
For example, where the database specifies literature from which data was extraced,
the CiTO ontology is used. The choice of the ontologies is driven by the use cases and
compatibility with other Linked Data projects, rather than formal ontological arguments.
We have not observed ontological contradictions based on these choices.

@@

The ongoing ChEMBL-RDF project is Open Source making it possible to others to convert ChEMBL to RDF too.
Technically, the choice of PHP makes it easy to share functionality between the
here described conversion scripts, and PHP-enabled web services, used to provide the Linked
Data interface to the ChEMBL-RDF data, as is fast enough for practical uses. These scripts do
not use a library for handling RDF data. While using such libraries increases the dependencies
for the scripts, they also ensure the generation of syntactically correct RDF triples. Continued
development of the ChEMBL-RDF scripts, however, has started using RDF libraries for this purpose
specificaly for triple generation of compound names, removing code in the scripts to escape
special characters, like double quotes.

>>>>

The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
The SQL data dump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
were then executed to generate RDF triples using a set of PHP scripts. The scripts use SQL to
query the database and output triples using print statements. Custom PHP scripts have been used
as a light-weight solution to provide a Linked Data API via a Apache webserver, as well as to
create data files with triples. These scripts are Open Source 
and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest
Group~\cite{Marshall2012}. This process has not changed much since its original description.

@@

However, where the database content is well-defined, specific ontologies have been used.
For example, where the database specifies literature from which data was extracted,
the CiTO ontology is used. The choice of the ontologies is driven by the use cases and
compatibility with other Linked Data projects, rather than formal ontological arguments.
We have not observed ontological contradictions based on these choices.

@@

The ongoing ChEMBL-RDF project is Open Source making it possible to others to convert ChEMBL to RDF too.
Technically, the choice of PHP makes it easy to share functionality between the
here described conversion scripts, and PHP-enabled web services, used to provide the Linked
Data interface to the ChEMBL-RDF data, as is fast enough for practical uses. These scripts do
not use a library for handling RDF data. While using such libraries increases the dependencies
for the scripts, they also ensure the generation of syntactically correct RDF triples. Continued
development of the ChEMBL-RDF scripts, however, has started using RDF libraries for this purpose
specifically for triple generation of compound names, removing code in the scripts to escape
special characters, like double quotes.
